### 2024-04-11 paper/paper.qmd
plots, figures, checks, models

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Hannah Yu
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(marginaleffects)
library(plotly)
library(tibble)
library(margins)
library(kableExtra)
library(arrow)
library(gridExtra)
library(cowplot)
library(lemon)
```

```{r}
ces2020 <-  read_parquet(here::here("data/analysis_data/cleaned_ces2020.parquet"))
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
```{r}
#| label: tbl-clean-data
#| tbl-cap: "Preview of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

ces2020[1:5, 1:11] |>
  kable()
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

ces2020 |>
  summary() |>
  kable()
```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

I utilized a Bayesian logistic regression model to examine the relationship between voters' political preferences and their media use preferences.The model is formulated as follows:

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{ABC}_i + \beta_2 \times \mbox{CBS}_i + \beta_3 \times \mbox{NBC}_i + \beta_4 \times \mbox{CNN}_ + \beta_5 \times \mbox{Fox_News}_ + \beta_6 \times \mbox{MSNBC}_ + \beta_7 \times \mbox{PBS}_ + \beta_8 \times \mbox{Other}_ + \beta_9 \times \mbox{TV_type}_ + \beta_10 \times \mbox{Party}\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim \mbox{Normal}(0, 2.5) \\
\beta_6 &\sim \mbox{Normal}(0, 2.5) \\
\beta_7 &\sim \mbox{Normal}(0, 2.5) \\
\beta_8 &\sim \mbox{Normal}(0, 2.5) \\
\beta_9 &\sim \mbox{Normal}(0, 2.5) \\
\beta_10 &\sim \mbox{Normal}(0, 2.5)
\end{align}

In this model, $y_i$ represents the binary outcome variable indicating whether an individual voted Biden (as opposed to Trump). The probability of voting for the Biden ($\pi_i$) is modeled using a logistic link function (\text{logit}($\pi_i$)), which is a linear combination of the intercept ($\alpha$) and the coefficients ($\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$, $\beta_10$)) corresponding to the predictor variables media use networks, TV news type, and party affiliations, respectively. Media use networks are being represented by the following predictor variables: \text{ABC}_i, \text{CBS}_i, \text{NBC}_i, \text{CNN}_i, \text{Fox_News}_i \text{MSNBC}_i ,\text{PBS}_i, and \text{Other}_i. The rest of the predictor variables are denoted as \text{TV_type}_i and \text{Party}_i, where $i$ indexes the individuals in the dataset.

The intercept and coefficients are assigned informative prior distributions to regularize the model. Specifically, a normal distribution with a mean of 0 and a standard deviation of 2.5 for each parameter is assumed.

This modeling approach is chosen for several reasons. Firstly, logistic regression is well-suited for binary outcome variables, making it appropriate for analyzing voting behavior. Additionally, Bayesian methods enables the incorporation of prior knowledge and uncertainty into our analysis, which provides more robust estimates of the model parameters. Alternative modeling approaches, such as linear regression models, were also considered. However, Bayesian logistic regression is chosen to fit the resulting binary variable of voter's decision.

We run the model in R [@citeR] using the `rstanarm` package of @citeRstanarm. We use the default priors from `rstanarm`. Rstanarm employs Markov chain Monte Carlo (MCMC) techniques to estimate the posterior distribution of the parameters. To avoid excessive runtime, we randomly sampled 2500 observations to fit the model. Model diagnostics, including convergence checks and posterior summaries, are available in the supplementary materials (see Appendix @sec-model-details).

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

```{r, fig.height = 10, fig.width = 10}

# Create individual plots for each news channel
plot_ABC <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(ABC))) +
  geom_bar(position = "dodge") +
  labs(title = "ABC", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_CBS <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(CBS))) +
  geom_bar(position = "dodge") +
  labs(title = "CBS", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_NBC <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(NBC))) +
  geom_bar(position = "dodge") +
  labs(title = "NBC", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_CNN <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(CNN))) +
  geom_bar(position = "dodge") +
  labs(title = "CNN", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_Fox_News <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(Fox_News))) +
  geom_bar(position = "dodge") +
  labs(title = "Fox News", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
#  facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_MSNBC <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(MSNBC))) +
  geom_bar(position = "dodge") +
  labs(title = "MSNBC", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
#  facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_PBS <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(PBS))) +
  geom_bar(position = "dodge") +
  labs(title = "PBS", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal()

plot_Other <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(Other))) +
  geom_bar(position = "dodge") +
  labs(title = "PBS", x = "Voted for Biden (1) or Trump (0)", y = "Count") +
  theme_minimal() +
  guides(fill = FALSE)

legend <- g_legend(plot_PBS + theme(legend.position='bottom'))

grid.arrange(plot_ABC, plot_CBS, plot_NBC, plot_CNN, plot_Fox_News, plot_MSNBC, plot_PBS + theme(legend.position ='none'), plot_Other, legend, nrow = 3)

```

```{r, fig.height = 6.6, fig.width = 10}

# Create individual plots for each news channel
plot_post <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(post))) +
  geom_bar(position = "dodge") +
  labs(title = "Posted", x = "Vote", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(ylim = c(0, 4000)) +  
  guides(fill = FALSE)

plot_comment <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(comment))) +
  geom_bar(position = "dodge") +
  labs(title = "Commented", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 4000)) +  
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_read <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(read))) +
  geom_bar(position = "dodge") +
  labs(title = "Read", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 4000)) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_follow <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(follow))) +
  geom_bar(position = "dodge") +
  labs(title = "Followed", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 4000)) +
  theme(plot.title = element_text(hjust = 0.5)) +
 # facet_wrap(~ ., ncol = 4) +
  guides(fill = FALSE)

plot_forward <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(forward))) +
  geom_bar(position = "dodge") +
  labs(title = "Forwarded", x = "Vote", y = "Count") +
  coord_cartesian(ylim = c(0, 4000)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5))

legend <- g_legend(plot_forward + theme(legend.position='bottom', 
                                        legend.text = element_text(size = 15, face="bold"), 
                                        legend.title = element_blank()))

grid.arrange(plot_post, plot_comment, plot_read, plot_follow, plot_forward + theme(legend.position ='none'), legend, nrow = 2)

```

```{r}
# Create individual plots for each news channel
plot_post <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(post, levels = c("not_selected", "selected")))) +
  geom_bar(position = "stack", stat = "count") +
  labs(title = "Posted", x = "Vote", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(ylim = c(0, 6000)) +  
  guides(fill = FALSE)

plot_comment <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(comment, levels = c("not_selected", "selected")))) +
  geom_bar(position = "stack", stat = "count") +
  labs(title = "Commented", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 6000)) +  
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) +
  guides(fill = FALSE)

plot_read <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(read, levels = c("not_selected", "selected")))) +
  geom_bar(position = "stack", stat = "count") +
  labs(title = "Read", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 6000)) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) +
  guides(fill = FALSE)

plot_follow <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(follow, levels = c("not_selected", "selected")))) +
  geom_bar(position = "stack", stat = "count") +
  labs(title = "Followed", x = "Vote", y = "Count") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 6000)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) +
  guides(fill = FALSE)

plot_forward <- ggplot(ces2020, aes(x = factor(voted_for), fill = factor(forward, levels = c("not_selected", "selected")))) +
  geom_bar(position = "stack", stat = "count") +
  labs(title = "Forwarded", x = "Vote", y = "Count") +
  coord_cartesian(ylim = c(0, 6000)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5)) 
 # facet_wrap(~ ., ncol = 4) +

#legend <- g_legend(plot_forward + theme(legend.position='bottom', 
                                        #legend.text = element_text(size = 15, face="bold"), 
                                       # legend.title = element_blank())

grid.arrange(plot_post, plot_comment, plot_read, plot_follow, plot_forward + theme(legend.position ='none'), nrow = 1)

```
```{r}
# Change labels for Fox_News variable
ces2020$Fox_News <- factor(ces2020$Fox_News, labels = c("No", "Yes"))

ggplot(ces2020, aes(x = Fox_News, fill = voted_for)) +
  geom_bar() +
  facet_wrap(facets = vars(Party)) +
  labs(title = "Relationship Between Fox News Consumption and Voting Behavior",
       x = "Fox News",
       y = "Count",
       color = "Voted for Biden (1) or Trump (0)") +
  scale_fill_brewer(palette = "Set1")
 # scale_color_manual(values = c("red", "blue")) +  # Customize colors for better visibility
  theme_minimal()
```
```{r}
# Change labels for Fox_News variable
ces2020$CNN <- factor(ces2020$CNN, labels = c("No", "Yes"))

ggplot(ces2020, aes(x = CNN, fill = voted_for)) +
  geom_bar() +
  facet_wrap(facets = vars(Party)) +
  labs(title = "Relationship Between CNN Consumption and Voting Behavior",
       x = "CNN",
       y = "Count",
       color = "Voted for Biden (1) or Trump (0)") +
  scale_fill_brewer(palette = "Set1")
 # scale_color_manual(values = c("red", "blue")) +  # Customize colors for better visibility
  theme_minimal()
```

```{r}
ces2020$TV_type <- factor(ces2020$TV_type, labels = c("National & Local", "National"))

scatter_plot <- ggplot(ces2020, aes(x = TV_type, fill = voted_for)) +
  geom_bar() +
  facet_wrap(facets = vars(Party)) +
  labs(title = "Relationship Between TV News Consumption and Voting Behavior",
       x = "TV Type",
       y = "Count",
       color = "Voted for Biden (1) or Trump (0)") +
  scale_fill_brewer(palette = "Set1")
 # scale_color_manual(values = c("red", "blue")) +  # Customize colors for better visibility
  theme_minimal()

# Show the plot
scatter_plot

```

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)
political_preferences <-
  readRDS(file = here::here("models/political_preferences.rds"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(
  list(
    "Support Biden" = political_preferences
  ),
  statistic = "mad"
  )
```

```{r}
# Load necessary libraries
library(car)

# Check for multicollinearity using VIF
vif(political_preferences)

# View the model summary
summary(political_preferences)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
 # theme_classic() +
 # theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
  #theme_minimal() +
 # scale_color_brewer(palette = "Set1") +
 # theme(legend.position = "bottom") +
 # coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

```{r}
#| label: fig-post_dist
#| fig-cap: Posterior distribution for logistic regression model 
#| echo: false
#| warning: false
#| message: false

pp_check(political_preferences) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-post_prior
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false

#### Read data ####
analysis_data1 <- read_parquet(here::here("data/analysis_data/cleaned_ces2020.parquet"))

# Convert variables to factors
#analysis_data$race <- factor(analysis_data$race)
#analysis_data$region <- factor(analysis_data$region)
#analysis_data$employ <- factor(analysis_data$employ)

# Create voted_for variable in binary form
analysis_data1$voted_for_binary <- ifelse(analysis_data1$voted_for == "Biden", 1, 0)

# Model 1 for n = 1000
set.seed(853)

ces2020_reduced <- 
  analysis_data1 |> 
  slice_sample(n = 2500)

posterior_vs_prior(political_preferences, size = 0.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=8)) +
  coord_flip()
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and race"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of race White", "Trace plot of race Black", "Trace plot of race Hispanic", "Trace plot of race Middle Eastern", "Trace plot of race Native American", "Trace plot of race Two or more races", "Trace plot of race Other"]
#| layout-ncol: 2

#plot(political_preferences1, "trace")
plot(political_preferences, "trace", "(Intercept)")

plot(political_preferences, "trace", "ABCYes")
plot(political_preferences, "trace", "CBSYes")
plot(political_preferences, "trace", "NBCYes")
plot(political_preferences, "trace", "CNNYes")
plot(political_preferences, "trace", "Fox_NewsYes")
plot(political_preferences, "trace", "MSNBCYes")
plot(political_preferences, "trace", "PBSYes")
plot(political_preferences, "trace", "OtherYes")

```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace2
#| fig-cap: "Trace plot of region"
#| fig-subcap: ["Trace plot of region Northeast", "Trace plot of region South", "Trace plot of region West"]
#| layout-ncol: 2

plot(political_preferences, "trace", "TV_typeNational Newscast")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace3
#| fig-cap: "Trace plot of region"
#| fig-subcap: ["Trace plot of region Northeast", "Trace plot of region South", "Trace plot of region West"]
#| layout-ncol: 2

plot(political_preferences, "trace", "PartyIndependent")
plot(political_preferences, "trace", "PartyOther")
plot(political_preferences, "trace", "PartyRepublican")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat plot"

plot(political_preferences, "rhat")
```

## Credibility Interval {#sec-credibility-interval}
@fig-modelresults1 shows the 90% credibility interval for the model.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"

modelplot(political_preferences, conf_level = 0.90, size = 0.2) +
  labs(x = "90% credibility interval")
```
\newpage

# References